Hybrid nonlinear system identification using neural networks

### 2020-11-24 systemID_NNs.tex
making progress in writing.

The following optimization problem is solved to compute the weights in the
neural network
\begin{align*}
\underset{W_i, b_i}{\textnormal{min}} \sum_{k=0}^{N_s-1} 
\dfrac{1}{2}(\hat{y}(k) - y(k))^2
 \end{align*}
in which $W_i$, $b_i$ are the weights in each hidden layer of the NN, $N_s$ is
the number of training samples, $\hat{y}$ are the predictions of the augmented
grey-box + neural network model, and $y$ are the measurements from the plant.
Note that this optimization requires an estimate of the initial grey-box state.
We assume that before the open-loop system identification experiment performed
for data collection, the plant is being operated at some steady-state such that
input to output responses of the plant and grey-box model are aligned with the
use of an integrating disturbance model. We can use the grey-box state at that
steady state as the initial condition for the above multi-step-ahead
prediction-error optimization.

@@

We examine the performance of the hybrid modeling framework
\eqref{eq:hybrid_discrete_time} in simulation studies. First, we generate
training data by simulating the plant model using a pseudo-random-binary-signal
(PRBS) manipulated input profile. Second, we train a black-box neural network
and the hybrid model \eqref{eq:hybrid_discrete_time} with network architectures
of [9, 16, 2] and [8, 16, 2]. Third, we examine trained model predictions and a
mean squared error (MSE) performance metric on validation data generated using
the plant model. Finally, we analyze the steady state cost curves obtained using
the trained models and their differences with the steady state cost 
curve of the plant model. These
simulations and analysis shed light on the data requirements of the hybrid and black-box
modeling frameworks, and enable an understanding of their performance if used
for real-time economic optimization in chemical plants.

\begin{figure}[!h]
  \centering
  \includegraphics[page=2, width=\textwidth]{tworeac_plots_nonlin.pdf}
  \caption{Model validation for the non linear plant}
  \label{fig:validation_nonlinear}
\end{figure}

\begin{figure}[!h]
  \centering
  \includegraphics[page=3, width=\textwidth]{tworeac_plots_nonlin.pdf}
  \caption{Steady state cost curve for the nonlinear plant, grey-box, black-box,
  residual, and hybrid model after training with 5 hours of data.}
  \label{fig:cost_nonlinear_5hours}
\end{figure}

\begin{figure}[!h]
  \centering
  \includegraphics[page=4, width=\textwidth]{tworeac_plots_nonlin.pdf}
  \caption{Steady state cost curve for the nonlinear plant, grey-box, black-box,
  residual, and hybrid model after training with 10 hours of data.}
  \label{fig:cost_nonlinear_10hours}
\end{figure}

%\begin{figure}[!h] \centering \includegraphics[page=1,
%  width=\textwidth]{tworeac_distmodels_nonlin.pdf} \caption{Steady state
%  grey-box states and disturbance estimates based on output and input
%  disturbance models} \label{fig:cost_nonlinear_10hours} \end{figure}

>>>>

The following optimization problem is solved to compute the weights in the
neural network
\begin{align*}
\underset{W_i, b_i}{\textnormal{min}} \sum_{k=0}^{N_s-1} 
\dfrac{1}{N_s}(\hat{y}(k) - y(k))^2
 \end{align*}
in which $W_i$, $b_i$ are the weights in each hidden layer of the NN, $N_s$ is
the number of training samples, $\hat{y}$ are the predictions of the augmented
grey-box + neural network model, $N_s$ is the number of training samples, and
$y$ are the measurements from the plant. Note that this optimization requires an
estimate of the initial grey-box state. We assume that before the open-loop
system identification experiment performed for data collection, the plant is
being operated at some steady-state such that input to output responses of the
plant and grey-box model are aligned with the use of an integrating disturbance
model. We can use the grey-box state at that steady state as the initial
condition for the above multi-step-ahead prediction-error optimization.

@@

We examine the performance of the hybrid modeling framework
\eqref{eq:hybrid_discrete_time} in simulation studies. First, we generate
training data by simulating the plant model using a pseudo-random-binary-signal
(PRBS) manipulated input profile. Second, we train a black-box neural network
and the hybrid model \eqref{eq:hybrid_discrete_time} with network architectures
of [9, 16, 2] and [8, 16, 2]. Third, we examine the trained model predictions
and a mean squared error (MSE) performance metric on validation data generated
using the plant model. Finally, we analyze the steady-state cost curves obtained
using the trained models and their differences with the steady-state cost curve
of the plant model. These simulations and analyses shed light on the data
requirements of the hybrid and black-box modeling frameworks, and enable an
understanding of their performance if used for real-time economic optimization
in chemical plants.

The Figure \ref{fig:validation_pred} shows predictions of the trained black-box
and hybrid models after training with 8 hours (480 samples) of data. We observe
that both the models can match the plant data accurately. The black-box model
shows slightly worse predictions in a few places. However, both the trained
models are adequate for real-time optimization in a feedback control system.

\begin{figure}[!h]
  \centering
  \includegraphics[page=2, width=\textwidth]{tworeac_plots_nonlin.pdf}
  \caption{Model validation for the nonlinear reactor model. Both the black-box
           and hybrid model can predict the plant responses accurately. The
           model predictions shown in this plot correspond to training with 8
           hours of data.}
  \label{fig:validation_pred}
\end{figure}

We examine the prediction accuracy of the models as a function of the number of
training samples using the mean squared error (MSE) metric computed as follows
on the validation data
\begin{align*}
  \textnormal{MSE} =  \sum_{k=0}^{N_s-1} 
  \dfrac{1}{N_s}(\hat{y}(k) - y(k))^2
\end{align*}
in which $\hat{y}$ are the model predictions and $y$ denotes the plant
measurements. The Figure \ref{fig:validation_mse} shows the variation of this
MSE metric as a function of the number of training data obtained with both the
black-box and hybrid models. We observe that the hybrid model provides better
performance than the black-box model in terms of overall prediction accuracy and
data requirements.

\begin{figure}[!h]
  \centering
  \includegraphics[page=9, width=\textwidth]{tworeac_plots_nonlin.pdf}
  \caption{Mean squared error (MSE) performance metric obtained with the
           black-box and hybrid model with increase in the number of training
           samples. We notice that the hybrid model slightly outperforms the
           black-box model in data requirements.}
  \label{fig:validation_mse}
\end{figure}

Next, we study the performance of the trained models in a steady-state economic
optimization problem, and compare the cost curves obtained with the trained
models with the cost curve of the plant model. The motivation is to understand
the performance of these models if used in real-time economic optimization
layers common in chemical process industries. The optimization problem solved is

\begin{align*}
  \underset{u_s}{\textnormal{min}} & \quad \ell(y_s, u_s) \\
  x_s &= f(x_s, u_s)\\ 
  y_s &= h(x_s)
\end{align*}
in which $x_s$, $y_s$, $u_s$ are the state, output, and control respectively at
steady-states. For the nonlinear reactor example, the cost is chosen as
$\ell(y_s, u_s) = c_1C_{Af} - c_2C_B$, which is a measure of the raw material
cost and product profit. The Figure  
\ref{fig:cost_curves}

\begin{figure}[!h]
  \centering
  \includegraphics[page=6, width=\textwidth]{tworeac_plots_nonlin.pdf}
  \caption{Steady-state cost curve for the nonlinear plant, grey-box, black-box,
  residual, and hybrid model after training with 5 hours of data.}
  \label{fig:cost_curves}
\end{figure}

\begin{figure}[!h]
  \centering
  \includegraphics[page=4, width=\textwidth]{tworeac_plots_nonlin.pdf}
  \caption{Steady-state cost curve for the nonlinear plant, grey-box, black-box,
  residual, and hybrid model after training with 10 hours of data.}
  \label{fig:cost_nonlinear_10hours}
\end{figure}

\section{Conclusion}
Hey.
%\begin{figure}[!h] \centering \includegraphics[page=1,
%  width=\textwidth]{tworeac_distmodels_nonlin.pdf} \caption{Steady state
%  grey-box states and disturbance estimates based on output and input
%  disturbance models} \label{fig:cost_nonlinear_10hours} \end{figure}
